The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-04-24 paper/ms.tex
a few more small revisions to section 2 in the new analytic derivation. more to come

We will check the formulae for the midplane offset against numerical tests
next. The effect of velocity offsets on actions is deferred to future work, as
we discuss in Section~\ref{ssec:lsr_var}.

\subsection{Numerical Methods} \label{ssec:action_comp}
We now move on to quantifying the argument made in Section~\ref{ssec:cartoon}.
We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
we use the torus-mapping technique first presented by
\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
calculate actions for an orbital time-series starting from a phase-space
position $(x, v)$ and integrated in a potential $\Phi$. We use the default
\texttt{MWPotential} as our potential, which is based on the Milky Way
potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
\citep{1997ApJ...490..493N}, and is fit to empirically match some
observations. We use the Dormand-Prince 8(5,3) integration scheme
\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.

>>>>

In the remainder of this section, we compare our analytic estimates of the effect on actions of a midplane offset against numerical calculations. The effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.

\subsection{Numerical Methods} \label{ssec:action_comp}
We now quantify the argument made in Section~\ref{ssec:cartoon} using numerical computations of the actions for a range of orbits in a model Galactic potential.
We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
we use the torus-mapping technique first presented by
\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
calculate actions for an orbital time-series starting from a phase-space
position $(x, v)$ and integrated in a potential $\Phi$. We use the default
\texttt{MWPotential} as our potential, which is based on the Milky Way
potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
\citep{1997ApJ...490..493N}, and is fit to empirically match some
observations. We use the Dormand-Prince 8(5,3) integration scheme
\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.
